The effects of probabilistic context inference on motor adaptation

### 2023-01-17 adaptation/article/article.tex
Some reviewrs' comments addressed.

\documentclass[a4paper,doc,floatsintext,natbib]{apa6}
%% \documentclass{article}
\usepackage[font=large]{caption}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{nameref}
\usepackage{todonotes}
\usepackage{authblk}
% \usepackage[nomarkers,figuresonly]{endfloat}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{lineno}
\linenumbers

@@

% For revision
\DeclareRobustCommand{\newcontent}[1]{#1}

@@

By introducing blocks of trials in which body dynamics are altered (e.g. a mechanical arm exerts a force on the participant's hand), experimenters are able to observe motor adaptation through the lens of motor error. Across many different motor adaptation experiments \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Davidson_Scaling_2004}, well-established phenomena have been observed: (i) the ability to recall previously-learned skills, called savings; (ii) the ability to return to unmodified dynamics, termed de-adaptation; (iii) the interference in motor learning between opposing manipulations in dynamics, called anterograde interference; (iv) spontaneous display of behavior consistent with a previously-learned adaptation, during trials where errors are forced to be zero, called spontaneous recovery.

To explain these phenomena, a number of computational models have been introduced, which adapt their motor commands after observing motor errors. The most studied are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, but Bayesian accounts have also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between forward models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.

@@

Recently, a new computational model for context-dependant motor learning based on Bayesian inference was introduced by \cite{Heald_Contextual_2021}, called COIN (for context inference). \cite{Heald_Contextual_2021} formalized context inference as a process that operates independently from motor learning, but is informed by it, establishing a loop whereby context inference also informs motor learning. With this model, \cite{Heald_Contextual_2021} showed that context inference causes the observed changes in the rate of motor learning in previous experiments \cite[e.g.][]{Herzfeld_Encoding_2018}.

In this work show that the process of context inference underlies more behavioral phenomena than previously shown. To do this, we used a minimal model for motor adaptation that includes context inference, which we derived by simplifying the COIN model (henceforth called sCOIN). We focused on the effects of uncertain contextual information on switching behavior, especially during so-called error-clamp trials, in which errors are forced to zero by experimenters. More specifically, we focused on the effects of perceptual noise, as well as feedback modalities, in context inference, which in turn affects behavior in ways that can be directly measured. We show that through context inference, switching behavior can display three main effects that have been previously attributed to hypothesized ad-hoc mechanisms: (1) The size of an adaptation dictates how quick and reliable switching between tasks is \citep{Oh_Minimizing_2019,Kim_Neural_2015}, which we explain in terms of the effects of perceptual noise on context inference. (2) Previously-learned adaptations can interfere with switching behavior \citep{Davidson_Scaling_2004}, which we explain in terms of uncertain context inference. (3) Training history (i.e. which adaptations have been learned and for how long) affects switching during error-clamp trials \citep{Vaswani_Decay_2013}, which we also attribute to uncertain context inference. To do this, we used the sCOIN model to simulate the experimental setups and the decision-making agents (i.e. participants) during those experiments.

@@

\subsection{Modeling context-dependent adaptation}
We focused on three main components of the COIN model: (1) context inference, (2) motor adaptation and (3) action selection. The processes defined by these component occur in this order, and each component informs the ones that follow.

@@

The COIN model contains, additionally to these three main components, components for learning new contexts (i.e. inferring the existence of a new context that had not been previously encountered by the agent), as well as the ability to infer subject-specific parameters such as a participant's assumed transition probabilities between contexts, which can differ from the real, hidden transition probabilities. Because we seeked to focus on switching behavior between previously-learned contexts, as well as in the perceptual aspect of context inference, we chose to fix the participant-inferred transition probabilities between contexts, as well as the total number of contexts; in our simulations, we assume that participants have already inferred the real values of these quantities. Because we mainly focus on switching behavior, as well as error-clamp trials (both of which involve alread-learned adaptations), these simplifications to the model have minimal effects on our results. See Methods for more details on the sCOIN model, as well as the differences between COIN and sCOIN.

@@

\begin{figure}
\centering
\includegraphics[]{./figures/figure_1.png}
\caption{Schematic representation of the model and illustrative simulations. (A) Inference done by the model at every trial. Clear circles represent the observables, i.e. motor commands $u_t$ and direct observations $y_t$ (e.g. cursor position). The true context $\zeta_t$ is not directly observable, but influences $y_t$. The dark rectangles represent the prior distributions for the inferred adaptation level $x_t$ (Normal-Gamma distribution) and the current context $\zeta_t$ (discrete distribution with known $\pi$; see Methods). At every trial, the context is inferred, then motor adaptation is carried out and finally a motor command is issued; the flow of this process is indicated by the gray arrow in the background, while black arrows show the direction of information flow. (B) Simulations obtained with the model in (A), using a simulated experimental setup similar to that by \cite{Davidson_Scaling_2004}, in which the context changes at trial 20 (vertical, dashed lines). A total of 2x2 experiments were simulated, with low and high levels of both cue uncertainty and observation noise. In the left panel, the states inferred by the model for each of the 2x2 simulated experiments, where each color represents one experiment, with the same colors as the panels on the right. On the right, each plot represents context inference $p(ctx)$ for one specific level of cue uncertainty and observation noise. The y-axis represents the posterior probability of each context $p(\zeta_t = j)$; the black line represents the baseline context (i.e. no adaptation), the colored line (with the same colors as the panel on the left) represents the only adaptation to be learned during the simulated experiment. The shaded areas represent the standard deviation around the mean, obtained across 50 simulated participants.}
\label{fig:model}
\end{figure}

\subsection{Contextual cues and feedback}
The behavioral phenomena which are the focus of this work can be explained as arising from the effects of contextual cues and sensory feedback provided to participants during the experiment. To illustrate these effects in a simple example, we first simulated a generic motor adaptation experiment similar to those performed by \cite{Davidson_Scaling_2004}, in which participants must adapt to the curl forces exerted by a mechanical arm on the participants' hands as they perform pointing movements.

The key to an intuitive understanding of the results presented below is to observe what happens when the presence and reliability of contextual cues is varied, as well as the perceptual noise on the position of the hand. In \fref{fig:model}B, a 2x2 grid of results is shown: each simulation in this grid is a combination of low or high contextual cue uncertainty (where high uncertainty is equivalent to presenting no cues), and low or high perceptual noise (representing how well participants can detect deviations from the straight-line movements).

@@

As we show below, these effects are at the heart of the behavioral phenomena observed in the experiments by \cite{Kim_Neural_2015}, \cite{Oh_Minimizing_2019}, \cite{Davidson_Scaling_2004}, and \cite{Vaswani_Decay_2013}, which we directly simulate in this work, alongside others that we discuss in the Discussion section.

@@

For clarity, we first introduce necessary terminology that is typically used in experimental studies. As an example, we will use a typical motor adaptation task in which participants have to make reaching movements while holding the handle of a mechanical arm that exerts a curl force on the participant's hand. Depending on the trial, the mechanical arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all. Let us define the baseline context $O$ as that in which the mechanical arm exerts no force. Contexts $A$ and $B$ can be defined as those with clockwise and counter-clockwise forces, respectively. Abusing notation, a usual statement is that $B = (-A)$, as the forces have the same magnitude but point in opposite directions. Similarly, one can define context $A/2$, with the same direction of adaptation as $A$, but half the magnitude. Finally, many experiments include a block of error-clamp trials at the end of the experiment, in which the mechanical arm forces the participant to make straight-line movements; we represent these with the letter $E$.

With this terminology, a typical experiment \cite[e.g.][]{Ethier_Spontaneous_2008} would have a block structure of $O-A-B-E$, or $O-A-(-A)-E$, which means that the participant goes through a block of trials with no external force applied ($O$), a number of trials with a clockwise curl force ($A$), a block with counter-clockwise forces ($B$), and finally a block with error-clamp trials ($E$). With repeated contexts \citep[e.g.][]{Oh_Minimizing_2019}, an experiment can be described as $O_1-A_1-O_2-\ldots$.

\subsubsection{Cue- and sensory feedback uncertainty affects switching behavior}
The term 'savings' refer to the ability to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in humans \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}. In an $O-A-O-A$ experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate than that observed during the first $A$ block. The related concept of quick de-adaptation occurs in $A-O$ transitions, where participants switch back to baseline without having to re-learn it.

@@

In \fref{fig:oh-2019}A we show the results of simulations with the model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results can be seen in the switches between contexts in the center panel, i.e. when the solid gray line (representing the true context) switches between 40, 0 and -40; in the two right-most panels, the thick black line, representing the agent's adaptation, quickly follows these switches.

\begin{figure}
\centering
\includegraphics{./figures/figure_2.png}
\caption{Switching between learned adaptations gated by context inference. Data from our simulations (first three columns) compared to data adapted from figure 2A by \cite{Kim_Neural_2015} and figure 4A from \cite{Oh_Minimizing_2019} (last column). Experimental and simulated data was, in all three experiments, averaged across all participants. In the first column, the simulated context inference is represented by the posterior probabilities over all available contexts. Each color (black, blue, green) represents a different context, with black always representing the baseline (i.e. no adaptation). Vertical, dashed lines represent switches in the real context. The inferred state (angle of the visuomotor rotation) is shown in the second column, with the same colors as in the first column. In the third and fourth columns, the black line represents the adaptation (i.e. response) displayed by the agent as a function of trial number. The thin gray lines represent the optimal adaptation, i.e. the size of the true visuomotor rotation during the task. (A) Experiment by \cite{Kim_Neural_2015}. Both experimental and simulated results are shown only up to trial 300, of the original 600. (B) An experiment by \cite{Oh_Minimizing_2019}.  Blue and black lines are as in A. (C) Same experiment as (B) but with a 10 degree adaptation.}
\label{fig:oh-2019}
\end{figure}

@@

In contrast, \cite{Oh_Minimizing_2019} explained the results of their second experiment by positing that when adaptations were small, participants did not identify this as a new context and opted instead for a modification of their baseline model (i.e. how they move normally). Under this single-context explanation, however, savings do not exist, and adaptations need to be learned anew ever time a context changes. On the other hand, savings are present under the sCOIN model, their manifestations being diminished by the slower context inference. \cite{Oh_Minimizing_2019} analyzed savings during their second experiment and found that savings do exist, although greatly diminished compared to the first experiment. This is in favor of the dual-context model, as we present it here. \cite{Oh_Minimizing_2019} showed a slow decay during error-clamp trials in their experiments, which they considered evidence for the single-context account of the second experiment. However, as we show below, this can also be explained in a dual-context model as an effect of slow context inference.

@@

\begin{figure}
\centering
\includegraphics{./figures/figure_3.png}
\caption{Motor error when switching back to a previously-learned adaptation. (A) Experimental results from \cite{Davidson_Scaling_2004} and simulations with the SCOIN model are shown. In the first column, each panel represents context inference for one group of participants (top: group A; bottom: group 3A), with each line representing the posterior probability of a context (black for the baseline O). The second column represents the error made by our simulated agent after returning to the previously-learned context, with blue and yellow representing groups -A and 3A, respectively. The last panel represents the same data, from the \cite{Davidson_Scaling_2004} experiment. All panels share the x-axis, representing the number of trials elapsed since the switch to the new context. All simulations were executed 32 times per group, to obtain a reliable mean; shaded areas represent the standard deviation across all simulations. (B) Simulated results for an experiment similar to \cite{Davidson_Scaling_2004}, but changing the contexts seen by the two groups from -A to -2A for the first group, and from 3A to 4A for the second group. The panels follow the same structure as (A), without the last panel for experimental results. (C) Simulations for an experiment in which the baseline context has been removed altogether.}
\label{fig:davidson-2004}
\end{figure}

To confirm this explanation, we simulated variants of the experiment in which the sCOIN model predicts that the difference between groups diminishes or disappears. First, in \fref{fig:davidson-2004}B, we simulated an experiment in which the contexts have more extreme adaptations, making them more different from baseline than in the \cite{Davidson_Scaling_2004} experiments. To do this, one group adapts in a $O-A-(-2A)$ paradigm, while the other group adapts in a $0-A-4A$ paradigm. As in the original experiment, the second contexts ($-2A$ for one group, $4A$ for the other) are equally spaced from the first context. However, given the larger distance from baseline, the baseline context has the same probability for both groups after the switch back to $A$. This change makes both simulated groups infer the correct context almost equally quickly, making the difference between their errors much smaller compared to the original experiment. Furthermore, we simulated an experiment with an identical structure to that of \cite{Davidson_Scaling_2004}, but eliminated the baseline context from the agent. The results can be seen in \fref{fig:davidson-2004}C, where the switches between contexts are made identically by the two groups.

@@

\cite{Vaswani_Decay_2013} also found participant-specific delays before the decay to baseline started after the error-clamp phase begun. As can be seen in \fref{fig:vaswani-2013}D, each simulated participant follows a different path of decay, with variations caused directly by perceptual noise. In our simulations, however, it is clear that no such systematic lag can be directly observed, which is most noticeable when looking at context inference (\fref{fig:vaswani-2013}C), which begins the switch as soon as the error-clamp trials begin. This can be further observed in \fref{fig:vaswani-2013}D, where we plot each simulated participant (one run of the simulations, color coded as before); because observation noise was chosen randomly for each run, some runs appear to contain a large delay before the decay begins. This falls in line with the experiments by \cite{Brennan_Decay_2015}, who found that the lag observed by \cite{Vaswani_Decay_2013} disappeared when controling for correlations in perceptual noise, as well as by using a balanced experimental design and unbiased analysis.

>>>>

\documentclass[a4paper,doc,floatsintext,natbib]{apa6}
%% \documentclass{article}
\usepackage[font=large]{caption}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{nameref}
\usepackage{todonotes}
\usepackage{authblk}
% \usepackage[nomarkers,figuresonly]{endfloat}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{lineno}
\usepackage{xcolor}  % For highlighting new content
\usepackage{soul}    % For highlighting new content
\linenumbers

@@

% For revision
\DeclareRobustCommand{\newcontent}[1]{\hl{#1}}

@@

By introducing blocks of trials in which body dynamics are altered (e.g. a mechanical arm exerts a force on the participant's hand), experimenters are able to observe motor adaptation through the lens of motor error. Across many different motor adaptation experiments \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Davidson_Scaling_2004}, well-established phenomena have been observed: (i) the ability to recall previously-learned skills \newcontent{upon re-exposure to a previous altered dynamic}, called savings; (ii) the ability to return to unmodified dynamics, termed de-adaptation; (iii) the interference in motor learning between opposing manipulations in dynamics, called anterograde interference; (iv) spontaneous display of behavior consistent with a previously-learned adaptation, during trials where errors are forced to be zero, called spontaneous recovery.

To explain these phenomena, a number of computational models have been introduced, \newcontent{in which motor commands are adapted based on observed motor errors}. The most well-studied models are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, \newcontent{while Bayesian accounts have also been shown to provide} an alternative explanation for savings and quick de-adaptation in the form of switching between forward models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.

@@

Recently, a new computational model for context-dependant motor learning based on Bayesian inference was introduced by \cite{Heald_Contextual_2021}, called COIN (for context inference). \cite{Heald_Contextual_2021} formalized context inference as a process that operates independently from motor learning, \newcontent{but both informs and is informed by it}. With this model, \cite{Heald_Contextual_2021} showed that context inference causes the observed changes in the rate of motor learning in previous experiments \cite[e.g.][]{Herzfeld_Encoding_2018}.

In this work we show that the process of context inference underlies more behavioral phenomena than previously shown. \newcontent{We believe that these phenomena stem directly from the process of active inference and the uncertainty that comes with it in a Bayesian setting, as done by} \cite{Heald_Contextual_2021}, \newcontent{and not from the specific of the generative model used. To show how this Bayesian approach accounts for these phenomena}, we used a minimal model for motor adaptation that includes context inference, which we derived by simplifying the COIN model (henceforth called sCOIN). We focused on the effects of uncertain contextual information on switching behavior, especially during error-clamp trials, in which errors are forced to zero by experimenters. More specifically, we focused on the effects of perceptual noise, as well as feedback modalities, in context inference, which in turn affects behavior in ways that can be directly measured. We show that through context inference, switching behavior can display three main effects that have been previously attributed to hypothesized ad-hoc mechanisms: (1) The size of an adaptation dictates how quick and reliable switching between tasks is \citep{Oh_Minimizing_2019,Kim_Neural_2015}, which we explain in terms of the effects of perceptual noise on context inference. (2) Previously-learned adaptations can interfere with switching behavior \citep{Davidson_Scaling_2004}, which we explain in terms of uncertain context inference. (3) Training history (i.e. which adaptations have been learned and for how long) affects switching during error-clamp trials \citep{Vaswani_Decay_2013}, which we also attribute to uncertain context inference. To do this, we used the sCOIN model to simulate the experimental setups and the decision-making agents (i.e. participants) during those experiments.

@@

\subsection{Experimental designs}
In the following sections, we discuss the experiments done by \cite{Kim_Neural_2015}, \cite{Oh_Minimizing_2019}, \cite{Davidson_Scaling_2004} and \cite{Vaswani_Decay_2013}. In all experiments, participants were asked to do reaching or shooting movements from a starting position to a known target, and the experimental manipulations can be clasified into two categories: (1) curl forces with mechanical arms, and (2) visuomotor rotations. In this section, we briefly describe these two manipulations and leave the details of the experimental design to the sections below where each study is discussed.

In the experiments with mechanical arms by \cite{Davidson_Scaling_2004,Vaswani_Decay_2013}, participants performed the movements while holding the handle of a mechanical arm; the arm exerted a force on the handle on each trial which depended on the speed of the participant's hand, creating a force that was perpendicular to the direction of movement, called a curl force (see \fref{fig:experiments}A). In these experiments, the participants' view of their hand and the mechanical arm was blocked, and they had to rely on a representation of their hand on a screen in front of them, which included the starting position and the target. The experiments by \cite{Davidson_Scaling_2004,Vaswani_Decay_2013} provided an accurate position of the participant's hand, with no noise added to it, and the cursor representing the hand is visible throughout the experiment; note that \cite{Vaswani_Decay_2013} added noise to the cursor on the screen during their third experiment, but only their first experiment is simulated here.

\begin{figure}
\centering
\includegraphics[]{./figures/experiments.png}
\caption{Mechanical arm and visuomotor rotation experiments. Green and blue circles represent the starting point and the target, respectively. Targets are placed on the circumference on both experimental paradigms. (A) Mechanical arm experiment. During adaptation trials, the force exerted by the arm on the handle pushes the hand away from the straight line. In error-clamp trials, the mechanical arm creates resistance against movements perpendicular to the straight line to the target. (B) Visuomotor rotation experiment. The correspondence between hand movements and the cursor is rotated (visuomotor rotation). The error is measured as the arc between the target and the place at which the trajectory (dotted line) crossed the circumference.}
\label{fig:experiments}
\end{figure}

When the mechanical arm was set to exert no force, movements made by human participants were close to straight lines between the starting position and the target. When the curl force was applied, movements deviated from this straight-line path, creating curved movements that nevertheless reached the target, as participants adjusted their trajectory before the end of the movement. As participants repeated the movements over many trials, they learned to adapt to these curl forces and went back to performing near straight-line movements. During error-clamp trials, the mechanical arm engaged a very stiff spring that caused participants' movements to be a straight line, regardless of the participants' applied forces (see \fref{fig:experiments}A). This allowed experimenters to present a zero-error feedback to the participant, while at the same time measuring the forces participants applied onto the spring.

Because participants could see the cursor, they adapted their movements in-flight, always reaching the target. Errors were measured as deviations from the straight line connecting starting and target positions.

In the experiments with visuomotor experiments by \cite{Kim_Neural_2015,Oh_Minimizing_2019}, participants moved a cursor on a screen by using a joystick \cite{Kim_Neural_2015}, or on the surface of a mirror using the pen of a digitizer pad \cite{Oh_Minimizing_2019}. Movements were done from a starting position in the middle of the screen to targets appearing on a circumference of 10cm in radius. The experimental manipulation took the form of a rotation between the hand movements and the cursor of the screen (see \fref{fig:experiments}B), centered on the starting position. A clockwise rotation of 20 degrees means that a hand movement forward translated into a movement towards the right on the screen or mirror. \cite{Kim_Neural_2015} used the convention that clockwise rotations are negative, while \cite{Oh_Minimizing_2019} used the opposite.

\cite{Oh_Minimizing_2019} added Gaussian noise to the visuomotor rotation at each trial. With this, each trial had a different visuomotor rotation, centered at 10 or 20 degrees, and with a standard deviation of 0.5 or 4 degrees, depending on the condition. In this work, we focused in the comparison between conditions 1a and 3, with means of 20 and 10 degrees, respectively, and standard deviation of 0.5 degrees.

In contrast to experiments with a mechanical arm, movements in visuomotor rotation experiments are always close to a straight line. Errors are measured in degrees, as the deviation between the target and the point in the 10cm circumference where the cursor crossed it (see \fref{fig:experiments}B).

\subsection{Modeling context-dependent adaptation}
We focused on three main components of the COIN model: (1) context inference, (2) motor adaptation and (3) action selection. The processes defined by these components occur in this order, and each component informs the ones that follow.

@@

The COIN model contains, additionally to these three main components, components for learning new contexts (i.e. inferring the existence of a new context that had not been previously encountered by the agent), as well as the ability to infer subject-specific parameters such as a participant's assumed transition probabilities between contexts, which can differ from the real, hidden transition probabilities. Because we seeked to focus on switching behavior between previously-learned contexts, as well as in the perceptual aspect of context inference, we chose to fix the participant-inferred transition probabilities between contexts, as well as the total number of contexts; in our simulations, we assume that participants have already inferred the real values of these quantities. Because we mainly focus on switching behavior, as well as error-clamp trials (both of which involve already-learned adaptations), these simplifications to the model have minimal effects on our results. See Methods for more details on the sCOIN model, as well as the differences between COIN and sCOIN.

@@

\begin{figure}
\centering
\includegraphics[]{./figures/figure_1.png}
\caption{Schematic representation of the sCOIN model and illustrative simulations. (A) Inference done by the model at every trial. Clear circles represent the observables, i.e. \newcontent{the information available to the model (and the brain), and used for adaptation}: motor commands $u_t$ and direct observations $y_t$ (e.g. cursor position). The true context $\zeta_t$ is not directly observable, but influences $y_t$. The dark rectangles represent the prior distributions for the inferred adaptation level $x_t$ (Normal-Gamma distribution) and the current context $\zeta_t$ (discrete distribution with known $\pi$; see Methods). At every trial, the context is inferred, then motor adaptation is carried out and finally a motor command is issued; the flow of this process is indicated by the gray arrow in the background, while black arrows show the direction of information flow. (B) Simulations obtained with the model in (A), using a simulated experimental setup similar to that by \cite{Davidson_Scaling_2004}, in which the context changes at trial 20 (vertical, dashed lines). A total of 2x2 experiments were simulated, with low and high levels of both cue uncertainty and observation noise. In the left panel, the states inferred by the model for each of the 2x2 simulated experiments, where each color represents one experiment, with the same colors as the panels on the right. On the right, each plot represents context inference $p(ctx)$ for one specific level of cue uncertainty and observation noise. The y-axis represents the posterior probability of each context $p(\zeta_t = j)$; the black line represents the baseline context (i.e. no adaptation), the colored line (with the same colors as the panel on the left) represents the only adaptation to be learned during the simulated experiment. The shaded areas represent the standard deviation around the mean, obtained across 50 simulated participants.}
\label{fig:model}
\end{figure}

\subsection{Contextual cues and feedback}
The behavioral phenomena which are the focus of this work can be explained as arising from the effects of contextual cues and sensory feedback provided to participants during the experiment. To illustrate these effects in a simple example, we first simulated a generic motor adaptation experiment similar to those performed by \cite{Davidson_Scaling_2004}: \newcontent{participants performed reaching movements from a central position to fixed targets on a circumference, then back to the central position. All targets were positioned on the horizontal plane, at shoulder height. Movements were performed holding a robotic arm which exerted velocity-dependent rotary forces, which, when ignored, would move the participant's hand away from the straight-line path to the target. See the following section for more details on the experiments by} \cite{Davidson_Scaling_2004}. The results of these simulations can be seen in \fref{fig:model}B.

The key to an intuitive understanding of the results presented below is to observe what happens when the presence and reliability of contextual cues is varied, as well as the perceptual noise on the position of the hand. \newcontent{To do this in our simulations, we added noise to the categorical variable that encodes contextual cues, and to the observations (see Methods for details), at two different levels (low and high noise). We simulated 50 participants, with 30 trials per participant}. In \fref{fig:model}B, a 2x2 grid of results is shown: each simulation in this grid is a combination of low or high contextual cue uncertainty (where high uncertainty is equivalent to presenting no cues), and low or high perceptual noise (representing how well participants can detect deviations from the straight-line movements).

@@

As we show below, these effects are at the heart of the behavioral phenomena observed in the experiments by \cite{Kim_Neural_2015}, \cite{Oh_Minimizing_2019}, \cite{Davidson_Scaling_2004}, and \cite{Vaswani_Decay_2013}, \newcontent{and our simulated data qualitatively matches their results, as well as} others that we discuss in the Discussion section.

@@

For clarity, we first introduce necessary terminology that is typically used in experimental studies. As an example, we will use a typical motor adaptation task in which participants have to make reaching movements while holding the handle of a mechanical arm that exerts a curl force on the participant's hand. Depending on the trial, the mechanical arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all. Let us define the baseline context $O$ as that in which the mechanical arm exerts no force. Contexts $A$ and $B$ can be defined as those with clockwise and counter-clockwise forces, respectively. Abusing notation, a usual statement is that $B = (-A)$, as the forces have the same magnitude but point in opposite directions. Similarly, one can define context $A/2$, with the same direction of adaptation as $A$, but half the magnitude. Finally, many experiments include a block of error-clamp trials at the end of the experiment, \newcontent{in which the mechanical arm forces the participant's hand into a straight-line trajectory towards the target by counteracting any lateral forces}, eliminating any error for the trial; we represent these with the letter $E$.

With this terminology, a typical experiment \cite[e.g.][]{Ethier_Spontaneous_2008} would have a block structure of $O-A-B-E$, or $O-A-(-A)-E$, which means that the participant goes through a block of trials with no external force applied ($O$), a number of trials with a clockwise curl force ($A$), a block with counter-clockwise forces ($B$), and finally a block with error-clamp trials ($E$). \newcontent{When participants are exposed to the same context multiple times }\citep[e.g.][]{Oh_Minimizing_2019}, an experiment can be described as $O_1-A_1-O_2-\ldots$, \newcontent{where $A_n$ is context $A$, presented to the participant for the n-th time.}

\subsubsection{Cue- and sensory feedback uncertainty affects switching behavior}
The term 'savings' refers to the ability to remember a previously-learned adaptation, \newcontent{observed as an instant recall of the previously-learned adaptation} \cite[e.g.][]{Kim_Neural_2015,Oh_Minimizing_2019}, or as accelerated re-learning \cite[e.g.][]{Kojima_Memory_2004}; in the case of instant recall, the phenomenon is often called 'switching'. Savings is almost universally observed in humans \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}. In an $O-A-O-A$ experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate than that observed during the first $A$ block. The related concept of quick de-adaptation occurs in $A-O$ transitions, where participants switch back to baseline without having to re-learn it.

@@

In \fref{fig:oh-2019}A we show the results of simulations with the model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results can be seen in the switches between contexts in \newcontent{the third and fourth columns}, where the solid gray line (representing the true context) switches between 40, 0 and -40, and the thick black line, representing the agent's adaptation, quickly follows these switches.

\begin{figure}
\centering
\includegraphics{./figures/figure_2.png}
\caption{Switching between learned adaptations gated by context inference. Data from our simulations (first three columns) compared to data adapted from figure 2A by \cite{Kim_Neural_2015} and figure 4A from \cite{Oh_Minimizing_2019} (last column). Experimental and simulated data was, in all three experiments, averaged across all participants (20, 11 and 11 participants, respectively). We simulated 1, 11 and 11 participants, respectively, as the simulations for \cite{Kim_Neural_2015} had almost no variability. In the first column, the simulated context inference is represented by the posterior probabilities over all available contexts. Each color (black, orange, green) represents a different context, with black always representing the baseline (i.e. no adaptation). Vertical, dashed lines represent switches in the real context. The inferred state (angle of the visuomotor rotation) is shown in the second column, with the same colors as in the first column. In the third and fourth columns, the black line represents the adaptation (i.e. response) displayed by the agent as a function of trial number. The thin gray lines represent the optimal adaptation, i.e. the size of the true visuomotor rotation during the task. (A) Experiment by \cite{Kim_Neural_2015}. Both experimental and simulated results are shown only up to trial 300, of the original 600. (B) An experiment by \cite{Oh_Minimizing_2019}.  Blue and black lines are as in A. (C) Same experiment as (B) but with a 10 degree adaptation.}
\label{fig:oh-2019}
\end{figure}

@@

In contrast, \cite{Oh_Minimizing_2019} explained the results of their second experiment by positing that when adaptations were small, participants did not identify this as a new context and opted instead for a modification of their baseline model (i.e. how they move normally). Under this single-context explanation, however, savings do not exist, and adaptations need to be learned anew every time a context changes. On the other hand, savings are present under the sCOIN model, their manifestations being diminished by the slower context inference. \cite{Oh_Minimizing_2019} analyzed savings during their second experiment and found that savings do exist, although greatly diminished compared to the first experiment. This is in favor of the dual-context model, as we present it here. \cite{Oh_Minimizing_2019} showed a slow decay during error-clamp trials in their experiments, which they considered evidence for the single-context account of the second experiment. However, as we show below, this can also be explained in a dual-context model as an effect of slow context inference.

@@

\begin{figure}
\centering
\includegraphics{./figures/figure_3.png}
\caption{Motor error when switching back to a previously-learned adaptation. (A) Experimental results from \cite{Davidson_Scaling_2004} and simulations with the sCOIN model are shown. In the first column, each panel represents context inference for one group of participants (top: group $-A$; bottom: group $3A$), with each line representing the posterior probability of a context (black for the baseline O). The second column represents the error made by our simulated agent after returning to the previously-learned context, with blue and yellow representing groups -A and 3A, respectively. The last panel represents the same data from eight participants per group (mean + SEM), from the \cite{Davidson_Scaling_2004} experiment. All panels share the x-axis, representing the number of trials elapsed since the switch to the new context. All simulations were executed 32 times per group, to obtain a reliable mean; shaded areas represent the SEM across all simulations. (B) Simulated results for an experiment similar to \cite{Davidson_Scaling_2004}, but changing the contexts seen by the two groups from -A to -2A for the first group, and from 3A to 4A for the second group. The panels follow the same structure as (A), without the last panel for experimental results. (C) Simulations for an experiment in which the baseline context has been removed altogether.}
\label{fig:davidson-2004}
\end{figure}

To confirm this explanation, we simulated variants of the experiment in which the sCOIN model predicts that the difference between groups diminishes or disappears. First, in \fref{fig:davidson-2004}B, we simulated an experiment in which the contexts have more extreme adaptations, making them more different from baseline than in the \cite{Davidson_Scaling_2004} experiments. To do this, one group adapts in a $O-A-(-2A)$ paradigm, while the other group adapts in a $O-A-4A$ paradigm. As in the original experiment, the second contexts ($-2A$ for one group, $4A$ for the other) are equally spaced from the first context. However, given the larger distance from baseline, the baseline context has the same probability for both groups after the switch back to $A$. This change makes both simulated groups infer the correct context almost equally quickly, making the difference between their errors much smaller compared to the original experiment. Furthermore, we simulated an experiment with an identical structure to that of \cite{Davidson_Scaling_2004}, but eliminated the baseline context from the agent. The results can be seen in \fref{fig:davidson-2004}C, where the switches between contexts are made identically by the two groups.

@@

\cite{Vaswani_Decay_2013} also found participant-specific delays before the decay to baseline started after the error-clamp phase began. As can be seen in \fref{fig:vaswani-2013}D, each simulated participant follows a different path of decay, with variations caused directly by perceptual noise. In our simulations, however, it is clear that no such systematic lag can be directly observed, which is most noticeable when looking at context inference (\fref{fig:vaswani-2013}C), which begins the switch as soon as the error-clamp trials begin. This can be further observed in \fref{fig:vaswani-2013}D, where we plot each simulated participant (one run of the simulations, color coded as before); because observation noise was chosen randomly for each run, some runs appear to contain a large delay before the decay begins. This falls in line with the experiments by \cite{Brennan_Decay_2015}, who found that the lag observed by \cite{Vaswani_Decay_2013} disappeared when controling for correlations in perceptual noise, as well as by using a balanced experimental design and unbiased analysis.
